### 2013-01-16 paper_sean/methods.tex
JMC final edits?

Each run had a population size of 200 and lasted 200 generations. We performed 20 HyperNEAT runs that differed only in the seed provided to the random number generator, which affected stochastic events such as mutation. To make a statistical comparison to the encoding from Glette et al., we conducted 19 runs using the original code to supplement the one run performed for that paper. 
Each gait was evaluated for fourteen seconds in reality, with interpolation from and to a stationary position in the first and last two seconds, respectively, as in Yosinski et al. effectively resulting in 12 seconds of full-speed
motion. In simulation gaits were evaluated for 12 seconds. 
All HyperNEAT parameters were identical to those in Yosinski et al. except for the frequency of the sine wave input to the ANN, which was lowered from 4.2Hz to 0.64Hz to reduce servo shutdowns. 
To further reduce servo shutdowns, we punished high-frequency gaits during evolution. We calculated the frequency of a gait as the average number of servo direction changes per leg per second. If this frequency was higher than the experimentally-selected threshold of 1.67 Hz, the measure of distance traveled by the center of mass during the gait was reduced exponentially by multiplying it by a discount factor of $e^{(1.67-\mathrm{freq})}$. Following Clune et al.~\cite{clune2009evolving}, the fitness equation was $2^{\mathrm{distance}^{2}}$. After evolving the gaits in simulation, the champion gait of the last generation of each of the 20 runs was transferred onto the real robot and the distance traveled was measured.

>>>>

Each run had a population size of 200 and lasted 200 generations. We performed 20 HyperNEAT runs that differed only in the seed provided to the random number generator, which affected stochastic events such as mutation. To make a statistical comparison to the encoding from Glette et al., we conducted 19 runs using the original code to supplement the one run performed for that paper. 
Each gait was evaluated for fourteen seconds in reality, with interpolation from and to a stationary position in the first and last two seconds, respectively, as in Yosinski et al., effectively resulting in 12 seconds of full-speed
motion. In simulation gaits were evaluated for 12 seconds. 
All HyperNEAT parameters were identical to those in Yosinski et al. except for the frequency of the sine wave input to the ANN, which was lowered from 4.2Hz to 0.64Hz to reduce servo shutdowns. 
To further reduce servo shutdowns, we punished high-frequency gaits during evolution. We calculated the frequency of a gait as the average number of servo direction changes per leg per second. If this frequency was higher than the experimentally-selected threshold of 1.67 Hz, the measure of distance traveled by the center of mass during the gait was reduced exponentially by multiplying it by a discount factor of $e^{(1.67-\mathrm{freq})}$. Following Clune et al.~\cite{clune2009evolving}, the fitness equation was $2^{\mathrm{distance}^{2}}$. After evolving the gaits in simulation, the champion gait of the last generation of each of the 20 runs was transferred onto the real robot and the distance traveled was measured.

### 2013-01-16 paper_sean/results.tex
JMC final edits?

The simulator likely improved performance because of the number of evaluations it enabled, both in terms of the population size (200 vs. 9) and the number of generations (200 vs. 20), leading
to a total difference of 40000 vs 180 per evolutionary run compared to Yosinski et al.~\cite{yos:clune}. Another
potential cause of improved performance is the lower noise in the simulator, which could have helped HyperNEAT find coordinated, regular, gaits, which perform better. On the physical robot, the noise in the evaluation was
substantial, preventing effective learning~\cite{yos:clune}. 
To investigate this hypothesis, we performed 20 runs in simulation with only 180 fitness evaluations, which was the number used in Yosinski et al.~\cite{yos:clune}. The simulated gaits performed
slightly, but not significantly, better than those evolved in hardware ($p = 0.1571$, mean fitness 7.9 cm/s).
Reduced noise thus may have had a small affect on performance, but the substantial performance gains that resulted from using a simulator likely came from the additional evaluations it afforded.
The encouragement of low-frequency gaits in this study
also may have aided performance, especially since in simulation the gaits were
high-frequency in a few early generations before rapidly settling to a
 range below the penalized threshold~(\figref{threeplots}-Right).

>>>>

The simulator likely improved performance because of the number of evaluations it enabled, both in terms of the population size (200 vs. 9) and the number of generations (200 vs. 20), leading
to a total difference of 40000 vs. 180 per evolutionary run compared to Yosinski et al.~\cite{yos:clune}. Another
potential cause of improved performance is the lower noise in the simulator, which could have helped HyperNEAT find coordinated, regular, gaits, which perform better. On the physical robot, the noise in the evaluation was
substantial, preventing effective learning~\cite{yos:clune}. 
To investigate this hypothesis, we performed 20 runs in simulation with only 180 fitness evaluations, which was the number used in Yosinski et al.~\cite{yos:clune}. The simulated gaits performed
slightly, but not significantly, better than those evolved in hardware ($p = 0.1571$, mean fitness 7.9 cm/s).
Reduced noise thus may have had a small affect on performance, but the substantial performance gains that resulted from using a simulator likely came from the additional evaluations the simulator afforded.
The encouragement of low-frequency gaits in this study
also may have aided performance, especially since in simulation the gaits were
high-frequency in a few early generations before rapidly settling to a
 range below the penalized threshold~(\figref{threeplots}-Right).

### 2013-01-16 paper_sean/simulator.tex
JMC final edits?

\subsubsection{Simulator:}
Gaits evolved in the simulator from Glette et al., which represents QuadraTot in the Nvidia PhysX physics engine\footnotemark[\value{footnote}], including the mass and size of the QuadraTot components and its degrees of freedom~(\figref{quadratotWhiteBooties}-Right). In the simulator, each individual joint range was limit as described above, but Smart Cropping was not included because we found that it hindered performance by limiting the types of gaits evolution could explore in early generations.

>>>>

\subsubsection{Simulator:}
Gaits evolved in the simulator from Glette et al.\footnotemark[\value{footnote}], which represents QuadraTot in the Nvidia PhysX physics engine, including the mass and size of the QuadraTot components and its degrees of freedom~(\figref{quadratotWhiteBooties}-Right). In the simulator, each individual joint range was limit as described above, but Smart Cropping was not included because we found that it hindered performance by limiting the types of gaits evolution could explore in early generations.
